---
title: "Proof of Concept"
author: "Wayne Monical"
date: "2024-11-12"
output: html_document
---


This document will prove that this project is doable. We will import and clean the elk migration data. We will create a static plot of the geographic elk migration data. We will create a dashboard to explore the geographic data, and publish it in a website.


## Import Data

[Data source](https://www.sciencebase.gov/catalog/item/5a9f2782e4b0b1c392e502ea)


```{r}
library(tidyverse)

library(ggmap)
```


Selecting on elk id, latitude, longitude, and date variables.
```{r}
file_path = 'raw_data/Elk GPS collar data from National Elk Refuge 2006-2015.csv'

elk = 
  read_csv(file_path) |> 
  janitor::clean_names() |> 
  select(
    elk_id,
    dt, 
    lat,
    long
  ) |> 
  mutate(
    year = year(dt),
    month = month(dt),
    day = day(dt),
    hour = hour(dt)
  )

```




## Create a static map plot

[using ggmap](https://www.nceas.ucsb.edu/sites/default/files/2020-04/ggmapCheatsheet.pdf)

[from stack exchange](https://stackoverflow.com/questions/33942186/using-r-and-ggplot-to-draw-separate-lines-between-gps-coordinates/33944974)




Filter for a single elk named bob. We'll use him for testing.
```{r}
bob = 
  elk |> 
  filter(
    elk_id == 572,
    year %in% c(2006, 2007)
  )
```


### Step 0: Stadia Maps API Key

[Sign up here](https://client.stadiamaps.com/signup/)

Register the API Key
```{r}
register_stadiamaps(key = '29074900-bb6e-4a71-8f91-454c28190f88',
                    write = FALSE)
```




### Step 1: Download map

Find the minimum and maximum latitude and longitude of Bob's journey. This will give us the range of map to download.

```{r}
min_lat = bob |> pull(lat) |> min()
max_lat = bob |> pull(lat) |> max()
rng_lat = abs(min_lat - max_lat)
lowerleftlat = min_lat 
upperrightlat = max_lat 


min_long = bob |> pull(long) |> min()
max_long = bob |> pull(long) |> max()
rng_long = abs(min_long - max_long)
lowerleftlon = min_long - rng_long 
upperrightlon = max_long + rng_long
```


Define area by `c(lowerleftlon, lowerleftlat, upperrightlon, upperrightlat)`

```{r}
 myLocation <- c(left = lowerleftlon,
                 bottom = lowerleftlat,
                 right = upperrightlon,
                 top = upperrightlat)
```


```{r}
myMap <- get_stadiamap(
  bbox=myLocation,
  maptype = "stamen_terrain",
  crop=FALSE)
```

### Step 2: Plot map with elk movement
```{r}
ggmap(myMap) +
geom_path(
  data = bob, 
  aes(x=long, y=lat, color = month))+
  geom_line(alpha = 0) +
  scale_color_gradientn(colours = rainbow(12))
```

## Create geographic dashboard


```{r}
library(leaflet)
```


```{r}
mean_lat = bob |> pull(lat) |> mean()
mean_long = bob |> pull(long) |> mean()

```


[from the leaflet Rstudio documentation](https://rstudio.github.io/leaflet/)

Basic usage
```{r}
m = leaflet() %>% 
  addTiles() %>%
  setView(mean_long, mean_lat, zoom = 13) %>%
  addPopups(mean_long, mean_lat, 'Here is the average lat and long of bob')
m
```


Adding markers
```{r}
leaflet(bob) %>%  addTiles() %>% addCircles(lng = ~long, lat = ~lat)
```


```{r}
samp_colors = sample(colors(distinct = FALSE), 12)
```




Adding colors to bob's months
```{r}
bob = bob |> 
  mutate(month_color = rainbow(12)[bob$month],
         real_colors = samp_colors[bob$month]
  )
```




Adding lines
```{r}
## I think this is what I want

# leaflet(bob) %>%  addTiles() %>% addPolylines(lng = ~long, lat = ~lat, color = ~real_colors)

## leaflet::addPolylines

```

